---
title: "Evaluation of the effects of changes in Smax and Srep on the shape of the Ricker recruitment curve"
output:
  bookdown::html_document2
  
---

```{r load, include=FALSE}

library(ggplot2)
library(cowplot)




sgenfun<-function(Sgen,theta,Smsy){
  a <- theta[1]
  b <- theta[2]
  prt <- Sgen * exp(a-b*Sgen)
  ss <- (Smsy-prt)^2
  return(ss=ss)

}


```

# The ricker curve

For the purpose of this demonstration we will assume that the salmon populations follow the deterministic Ricker curve withe the following specification.

$R_t = a*S_t*exp(-b*S_t)$ 

$S_{max} = \frac{1}{b}$

$S_{rep} = \frac{log(a)}{b}$

Where

$R_t$ is the recruitment for brood year t
$S_t$ is the spawners for brood year t
$a$ is the productivity parameter (slope near the origin)
$b$ is the inverse of the capacity parameter
$S_{max}$ is the spawners that maximize recruits (spawners at the peak of the curve)
$S_{rep}$ is the spawners at replacement, x spawners will produce x recruits

In the following sections we will investigate the effects of assuming time varying $a$ and $b$ and keeping $Srep$ constant or allowing it to vary. 


# Capacity changes

First we evaluate the scenario in which $b$ changes and Srep is allowed to vary according to the equations in the  previous sections.  When $Srep$ varies, the $a$ parameter is kept constant and $Srep$ increases as $Smax$ increases. In this scenario, higher values of $Smax$/$Srep$ lead to higher recruitment at the same $S_t$ values (see left panel on Figure \@ref(fig:bvary)). 

When $b$ varies but $Srep$ is kept constant, the $a$ parameter is adjusted to allow for the constant $Srep$. In this scenario, $a$ decreases as $Smax$ increases. This leads to high productivities for low values of $Smax$ (see left panel on Figure \@ref(fig:bvary)). CW note: I am not sure this second scenario would make sense from a biological perspective, which mechanism would lead to increasing productivity (improved survival) as $Smax$ declines? Maybe less competition?

When looking at the effects on the estimates of $Smsy$, we see that Smsy increases linearly with $Smax$ when $Srep$ varies , but increases asymptotically when $Srep$ is kept constant (and productivity varies) (Figure \@ref(fig:smsybvary)). $Sgen$ tends to increase linearly when Smax increases, but the increase has higher slope in the scenario where $Srep$is kept constant (Figure \@ref(fig:smsybvary)). 


```{r, echo=FALSE, warning=FALSE,message=FALSE ,results='hide'}




#if Smax and Srep varies

a<- 5
Smax <- seq(7000,14000,by=1000)
b <- 1/Smax
Srep <- log(a)/b

S <- seq(500,30000, by=500)


Sgen <- NULL
Smsy <- (1-gsl::lambert_W0(exp(1-log(a))))/b
SmsyHW <- log(a)/b*(.5-0.07*log(a))


R<-NULL 
for(i in seq_along(Smax)){

  Rt <- a*S*exp(-b[i]*S)

  R<-c(R,Rt)

  Sgen[i] <- optimize(sgenfun, interval = c(0, Smsy[i]),theta=c(log(a),b[i]),Smsy=Smsy[i])$minimum

}

dfp<-data.frame(S=rep(S,length(Smax)),
 R=R,
 Smax=rep(Smax,each=length(S)),
 Srep=rep(Srep,each=length(S)),
 Smsy=rep(Smsy,each=length(S)))

p <- ggplot(dfp) +
  geom_line(aes(x=S, y=R, color=as.factor(Smax)), size=1.3) +
  geom_vline(aes(xintercept=Srep, color=as.factor(Smax))) +
 # geom_vline(aes(xintercept=Smax, color=as.factor(Smax))) +
theme_bw(14)+ scale_color_viridis_d() +ggtitle("(a) Smax and Srep vary") +
coord_cartesian(ylim=c(0,28000))


#Smsy and smax
#po <- ggplot(dfp) +
#  geom_line(aes(x=S, y=R, color=as.factor(Smax)), size=1.3) +
#  geom_vline(aes(xintercept=Smsy, color=as.factor(Smax))) +
#  geom_vline(aes(xintercept=Smax, color=as.factor(Smax)), linetype = "dashed") +
#theme_bw(14)+ scale_color_viridis_d() +ggtitle("Smax and Srep varies")
#po



Smax2 <- seq(7000,14000,by=1000)
b2 <- 1/Smax2
Srep2 <- 16094.38
a2<- exp(Srep2*b2) 


Smsy2 <- (1-gsl::lambert_W0(exp(1-log(a2))))/b2
SmsyHW2 <- log(a2)/b2*(.5-0.07*log(a2))

Sgen2 <-NULL
S <- seq(500,30000, by=500)

R<-NULL 
for(i in seq_along(Smax2)){

  Rt <- a2[i]*S*exp(-b2[i]*S)

  R<-c(R,Rt)

  Sgen2[i] <- optimize(sgenfun, interval = c(0, Smsy2[i]),theta=c(log(a2[i]),b2[i]),Smsy=Smsy2[i])$minimum

}

dfp2<-data.frame(S=rep(S,length(Smax2)),
 R=R,
 Smax=rep(Smax2,each=length(S)),
 Srep=rep(Srep2,each=length(S)))

p2 <- ggplot(dfp2) +
  geom_line(aes(x=S, y=R, color=as.factor(Smax)), size=1.3) +
  geom_vline(aes(xintercept=Srep, color=as.factor(Smax))) +
 # geom_vline(aes(xintercept=Smax, color=as.factor(Smax))) +
theme_bw(14)+ scale_color_viridis_d() +ggtitle("(b) Smax varies/ Srep constant") +
coord_cartesian(ylim=c(0,28000))







```

```{r bvary, echo=FALSE, fig.height = 5, fig.width = 12, fig.cap = "Potential shapes of the Ricker curve when b varies and Srep varies (left) or is kept constant (right). Vertical lines indicate Srep values."}
plot_grid(p, p2)

```


```{r smsybvary, echo=FALSE, fig.height = 5, fig.width = 12, fig.cap = "Increases in Smsy i relation to Smax when Srep varies or is kept constant."}

df<-data.frame(Smsy=c(Smsy,Smsy2),
  Smax=c(Smax,Smax2),
  a=c(rep(a,length(Smax)),a2),
 Srep=c(Srep,rep(Srep2,length(Smax))),
 Sgen=c(Sgen,Sgen2),
 scenario=rep(c("Srep_varies","Srep_constant"),each=length(Smax)))

pb <- ggplot(df) +
   geom_line(aes(x=Smax, y=Smsy,color=scenario), size=1.3) +
 theme_bw(14) + scale_color_viridis_d() +ggtitle("Smax")

 pbg <- ggplot(df) +
   geom_line(aes(x=Smax, y=Sgen,color=scenario), size=1.3) +
 theme_bw(14) + scale_color_viridis_d()  +ggtitle("Sgen")


plot_grid(pb,pbg)



```

# Productivity changes

In the scenarios when the productivity parameters are driving change, a similar exercise can be done. In this case the choice between $Smax$ and $Srep$ being fixed depends on biological mechanism: whether we think that recruitment is reduced at most levels of spawner abundances ($Smax$ if fixed) or we think the spawner abundance where density dependence result in over compensation starts at a lower spawner abundance ($Srep$ is fixed).

In the first scenario, productivity ($a$) changes and $Smax$ is kept constant, higher productivities lead to higher recruitment at all levels of spawner abundances (see left panel on Figure \@ref(fig:avary)). The $Srep$ value increases as productivity increases. 

In the second scenario, productivity ($a$) changes and $Srep$ is kept constant, therefore $Smax$ is allowed to vary. In this case lower productivities will have higher $Smax$ (inverse relation) (see right panel on Figure \@ref(fig:avary)). 

The two scenarios differ regarding the impacts on $Smsy$. When $Srep$ is allowed to vary and $Smax$ is kept constant, $Smsy$ increases as productivity increases. On the other hand, when $Srep$ is kept constant, $Smsy$ declines as productivity increases. $Sgen$ declines as a increases for both scenarios, but declines are sharper when $Sgen$

```{r, echo=FALSE, warning=FALSE,message=FALSE ,results='hide'}

#if b and Srep varies

a3<- seq(3,8,by=.5)
Smax3 <- 10000
b3 <- 1/Smax3
Srep3 <- log(a3)/b3

Smsy3 <- (1-gsl::lambert_W0(exp(1-log(a3))))/b3


Sgen3<-NULL

S <- seq(500,30000, by=500)

R<-NULL 
for(i in seq_along(a3)){

  Rt <- a3[i]*S*exp(-b3*S)

  R<-c(R,Rt)
  Sgen3[i] <- optimize(sgenfun, interval = c(0, Smsy3[1]),theta=c(log(a3[i]),b3),Smsy=Smsy3[i])$minimum
}

dfp<-data.frame(S=rep(S,length(a3)),
 R=R,
 a=rep(a3,each=length(S)),
 Srep=rep(Srep3,each=length(S)))

p3 <- ggplot(dfp) +
  geom_line(aes(x=S, y=R, color=as.factor(a)), size=1.3) +
  geom_vline(aes(xintercept=Srep, color=as.factor(a))) +
 # geom_vline(aes(xintercept=Smax, color=as.factor(Smax))) +
theme_bw(14)+ scale_color_viridis_d() +ggtitle("a and Srep vary")+
coord_cartesian(ylim=c(0,32000))



#if Srep is kept constant

a4<- a3 
Srep4 <- 16094.38
b4 <- log(a4)/Srep4
Smax4 <- 1/b4

Sgen4 <-NULL
S <- seq(500,30000, by=500)
Smsy4 <- (1-gsl::lambert_W0(exp(1-log(a4))))/b4
Smsy4HW<-log(a4)/b4*(.5-0.07*log(a4))
R<-NULL 
for(i in seq_along(a4)){

  Rt <- a4[i]*S*exp(-b4[i]*S)

  R<-c(R,Rt)
  Sgen4[i] <- optimize(sgenfun, interval = c(0, Smsy4[1]),theta=c(log(a4[i]),b4[i]),Smsy=Smsy4[i])$minimum
}

dfp4<-data.frame(S=rep(S,length(Smax4)),
 R=R,
 Smax=rep(Smax4,each=length(S)),
 a=rep(a4,each=length(S)),
 Srep=rep(Srep4,each=length(S)))

p4 <- ggplot(dfp4) +
  geom_line(aes(x=S, y=R, color=as.factor(a)), size=1.3) +
  geom_vline(aes(xintercept=Srep, color=as.factor(a))) +
 # geom_vline(aes(xintercept=Smax, color=as.factor(Smax))) +
theme_bw(14)+ scale_color_viridis_d() +ggtitle("a varies/ Srep constant")+
coord_cartesian(ylim=c(0,32000))







```

```{r avary, echo=FALSE, fig.height = 5, fig.width = 12, fig.cap = "Potential shapes of the Ricker curve when a varies and Srep varies (left) or is kept constant. Vertical lines indicate Srep values."}
plot_grid(p3, p4)

```

```{r smsyavary, echo=FALSE, fig.height = 5, fig.width = 12, fig.cap = "Increases in Smsy i relation to Smax when Srep varies or is kept constant."}

df<-data.frame(Smsy=c(Smsy3,Smsy4),
  Smax=c(rep(Smax3,length(a3)),Smax4),
  a=c(a3,a4),
 Srep=c(Srep3,rep(Srep4,length(a4))),
 Sgen=c(Sgen3,Sgen4),
 scenario=rep(c("Srep_varies","Srep_constant"),each=length(a3)))

pb2 <- ggplot(df) +
   geom_line(aes(x=a, y=Smsy,color=scenario), size=1.3) +
 theme_bw(14) + scale_color_viridis_d() + ggtitle("Smsy")



pbg2 <- ggplot(df) +
   geom_line(aes(x=a, y=Sgen,color=scenario), size=1.3) +
 theme_bw(14) + scale_color_viridis_d() + ggtitle("Sgen")



plot_grid(pb2,pbg2)

```